### 2017-11-11 TeX/Preface.tex

\documentclass[./Research.tex]{subfiles}

\begin{document}

	The purpose of our research is to understand how Artificial Neural Networks can support multiple classes of computations. This document will contain notes, diagrams, and explantations pertaining to each of the experiments we will conduct to investigate this problem. Code for each subsection of this document can be found at \url{https://github.com/DariusBxsci/NeuralNetworkResearch/tree/master/NeuralNets}.

\end{document}