### 2017-11-21 TeX/NeuralNetwork_3.tex
added some notes for lollipop model

\documentclass[./Research.tex]{subfiles}

\begin{document}

\def\layersep{2.5cm}

	When constructing Neural Networks, we generally consider the order of inputs to neurons or the order of inputs in a layer, to be irrelivent. However, there is no reason that the order of inputs should be self-evident, and in many cases, the best order of inputs may not be known at all. To examine this, let us create a simple model for a single neuron, where it accepts any number of inputs and has a single output which is the result of applying the 'eval' function to the inputs.

% diagram of neuron

	Each input is a tuple where the first index is either an operator or operand, and the second index contains the input's position in the eval function.

% diagram of structure of each input

\end{document}