### 2017-11-26 TeX/Preface.tex

\documentclass[./Research.tex]{subfiles}

\begin{document}

    \begin{enumerate}
        \item What is symbolic computation?
        \item What is calculation?
        \item What is meant by a computational class?
    \end{enumerate}

     Code for each subsection of this document can be found at \url{https://github.com/DariusBxsci/NeuralNetworkResearch/tree/master/NeuralNets}.

\end{document}