### 2017-11-26 TeX/Preface.tex

\documentclass[./Research.tex]{subfiles}

\begin{document}

    \begin{enumerate}
        \item What is symbolic computation?

        A symbolic computation is a calculation performed with symbolic representations of values and operations. A simple example would be the expression $(x + 1)(x - 1)$ which would evaluate to $x^2 - 1$, rather than to some numerical result.

        \item What is calculation?

        A calculation is a process by which one or more inputs is transformed into one or more results. One may calculate that the product of 5 and 4 is 20.

        \item What is meant by a computational class? \textcolor{red}{- do you mean complexity class?}

    \end{enumerate}

     Code for each subsection of this document can be found at \url{https://github.com/DariusBxsci/NeuralNetworkResearch/tree/master/NeuralNets}.

\end{document}