### 2017-12-02 TeX/manuscript.tex
Added to manuscript.tex and wrote in ResearchNotes about a possible direction for research

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{setspace, fancyhdr, hyperref, xcolor, tikz}
\usetikzlibrary{positioning, intersections, calc, arrows}
\usepackage[square,numbers]{natbib}
\bibliographystyle{abbrvnat}

\newcommand{\thetitle}{Symbolic Manipulation and Computation in the Same Graph}

\pagestyle{fancy}
\fancyhf{}
\rhead{\thetitle}
\lhead{Darius Barbano}
\rfoot{Page \thepage}
\doublespacing

\author{Darius Barbano}
\title{\thetitle}
\date{}

\begin{document}

\maketitle
\newpage
\tableofcontents
\newpage

\begin{abstract}
    General artificial intelligence refers to machine intelligence than performs a task as successfully as a human does. A fundamental difference between human neural network and current machine neural networks is that only human networks combine symbolic reasoning with computation. Here we show how a graph can \ldots
\end{abstract}
\newpage

\section{Background}

    \begin{enumerate}
        \item What is symbolic computation?

        A symbolic computation is a calculation performed with symbolic representations of values and operations. A simple example would be the expression $(x + 1)(x - 1)$ which would evaluate to $x^2 - 1$, rather than to some numerical result.

        \item What is calculation?

        A calculation is a process by which one or more inputs is transformed into one or more results. One may calculate that the product of 5 and 4 is 20.

        \item What is meant by a computational class? \textcolor{red}{- do you mean complexity class?}

    \end{enumerate}

     Code for each subsection of this document can be found at \href{https://github.com/DariusBxsci/NeuralNetworkResearch/tree/master/NeuralNets}{NeuralNetworkResearch}.

\section{Methods}

\subsection{Network Construction}

 Figure \ref{fig:sig-neuron} illustrates a neuron that receives three ordered inputs.

 \begin{figure}[h]
 \centering
 \begin{tikzpicture}

    \node[circle, draw=black] (soma) {Soma};
    \node[below= 2cm of soma] (dendrite) {};

    \begin{scope}[node distance=1mm and 10mm]
    \node[below right = of soma] (operator) {operator};
    \node[below = of operator] (operand1) {operand1};
    \node[below = of operand1] (operand2) {operand2};
    \end{scope}

    \draw[->] (dendrite) -> (soma);

    %nodes along dendrite
    \node (operator'dendrite) at (intersection of operator--soma|-operator and dendrite--soma){};

    \node (operand1'dendrite) at (intersection of operand1--soma|-operand1 and dendrite--soma){};

    \node (operand2'dendrite) at (intersection of operand2--soma|-operand2 and dendrite--soma){};

    \draw[->] (operator) -- (operator'dendrite) node[midway, fill=white] {1};
    \draw[->] (operand1) -- (operand1'dendrite) node[midway, fill=white] {2};
    \draw[->] (operand2) -- (operand2'dendrite) node[midway, fill=white] {3};

 \end{tikzpicture}
 \caption{Single neuron receiving ordered intputs. \label{fig:sig-neuron}}
\end{figure}

While this model is suitable for biological networks of neurons, a neural network in the computational sense would look more like an abstract syntax tree. In the following diagram, a network of operators and operands form a tree-like structure in which elements at lower levels represent values to manipulate which at higher levels interact with each other through arithmetic operations to produce a result. Any computation which involves a combination of operations on one or more values can be represented in this structure, such as logical operations or large-scale machine learning models.

\newpage

 \begin{figure}
 \centering
 \begin{tikzpicture}

\tikzstyle{bordered} = [shape=rectangle, rounded corners, draw, outer sep=0,inner sep=6,minimum size=15]
\tikzstyle{con} = [thick,->]
\tikzstyle{message} = [shape=rectangle, thick, minimum width=2cm]

	\node [bordered,red] (v1) at (1.5,8) {3};
	\node [bordered,red] (v5) at (3.5,8) {4};
	\node [bordered,red] (v4) at (4.5,8) {4};
	\node [bordered,red] (v3) at (6.5,8) {2};

	\node [bordered,blue] (v6) at (5.5,9) {$\div$};
	\node [bordered,blue] (v2) at (2.5,9) {$-$};

	\node [bordered,blue] (v7) at (4,9.5) {$*$};

	\node [bordered,black] (v8) at (4,11) {eval()};

	\node [black] (v9) at (4,12.5) {\Huge -2};

	\draw [con] (v1) edge (v2);
	\draw [con] (v3) edge (v6);
	\draw [con] (v4) edge (v6);
	\draw [con] (v5) edge (v2);

	\draw [con] (v2) edge (v7);
	\draw [con] (v6) edge (v7);

	\draw [con] (v7) edge (v8);

	\draw [con] (v8) edge (v9);

 \end{tikzpicture}
 \caption{An Abstract Syntax Tree of arithmetic operations. \label{fig:sig-neuron}}
\end{figure}

\section{Results}

\section{Conclusions}

\bibliography{references}

\end{document}